---
output: html_document
params:
  mlrassoarea:
    label: "Soil Survey Area Symbol:"
    value: "11-GAL"
  fiscalyear:
    label: "Fiscal Year"
    value: "2020"
---

# Project Impact Score

## Concept

Measuring the impact of a project can evaluated using 5 different metrics:

*	\# of Area Symbols which contain the polygons affected (N)
*	\# of National Mapunit Symbols altered (M)
*	\# of Polygons changed (P)
*	\# of Acres updated (A)
*	\# of people in updated area (H)

Using the factors together as a contribution to the total impact, a simple equation can be represented as shown:

$$N+M+P+A+H=I$$

Each of the above values from a single project must compared to the total amount in existence within the Soil Survey Office Area to determine a relative impact across the area.  Then each value is multiplied by a factor of 20 to normalize the values on a scale from 0 - 100.  The zero value representing a project with the least impact and 100 the most.  See equation below:

$$\left\{ \left( N_\rho\over \sum N \right) \times 20 \right\} + \left\{ \left( M_\rho\over \sum M \right) \times 20 \right\} + \left\{ \left( P_\rho\over \sum P \right) \times 20 \right\} + \left\{ \left( A_\rho\over \sum A \right) \times 20 \right\} + \left\{ \left( H_\rho\over \sum H \right) \times 20 \right\} = I$$


## Examples

Let's look at some examples:

### Example #1 (Clinton, Missouri)

MLRA 112 - Zaar silty clay, 1 to 3 percent slopes

$$\left\{ \left( 9\over 61 \right) \times 20 \right\} + \left\{ \left( 2\over 2841 \right) \times 20 \right\} + \left\{ \left( 2193\over 261662 \right) \times 20 \right\} + \left\{ \left( 126454\over 14686292 \right) \times 20 \right\} + \left\{ \left( 156388\over 3910405 \right) \times 20 \right\} = 4.10$$

Results for each factor:

$$\{2.95\}+\{0.01\}+\{0.17\}+\{0.17\}+\{0.80\}=4.10$$

As you can see, the number of area symbols in a project greatly influences the result.

### Example #2 (Union, Missouri)

MLRA 115B - Alluvial map units and hydric classifications within St Louis

$$\left\{ \left( 7\over 48 \right) \times 20 \right\} + \left\{ \left( 3\over 2976 \right) \times 20 \right\} + \left\{ \left( 241\over 189620 \right) \times 20 \right\} + \left\{ \left( 46776\over 8899209 \right) \times 20 \right\} + \left\{ \left( 4584762\over 5022683 \right) \times 20 \right\} = 21.32$$

Results of each factor:

$$\{2.92\}+\{0.02\}+\{0.03\}+\{0.11\}  +\{18.26\}=21.32$$

## Discussion

What can we use a project impact score for?  The PIS can be used as a tool to help with ranking the importance of a project.  It could also be useful for justifying a lower acreage project.  Another potential use for the PIS is for region-wide planning.  A high PIS means more SDQS workload.  The PIS could also help keep projects within an office balanced with equal high and low impact projects addressed each year.

There are several drawbacks with this method of determining impact.  The most obvious is the tendency for each metric to be related (If you have every acre of your whole area in a project, you'll obviously also have every polygon, every mapunit and every area symbol in it as well).  So in theory, a project with these values would have a PIS of 100.  On the other hand you could have a very small acreage project with at least one polygon in each area symbol and end up with a PIS of just over 20.  

It is also important to note, there are many other metrics which could be (and maybe should be) included in this calculation, such as amount of soils data updated, the egregiousness of the errors fixed, land use changes and so on.  However, the metrics included in this formula were easy to obtain and although not comprehensive, has value in understanding and comparing projects throughout the region.  

My recommendations for Region-wide planning:  Encourage at least one MLRA project each year in each office with a PIS of ~10.  Encourage ~50% of EVAL projects to have PIS > 0.50.  This is just a general recommendation and expert judgement will also be needed on a case by case basis.

## Data Output

This data below is from `r params$mlrassoarea` and represents projects in `r params$fiscalyear`.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r load libraries}
#### THIS SCRIPT IS FOR GENERATING PROJECT IMPACT SCORES FOR AN INDIVIDUAL SSO IN REGION 11 FOR A GIVEN FISCAL YEAR

# Load libraries
library(soilDB)
library(sp)
library(plyr)
library(sf)
library(dplyr)
library(stringr)
library(cdlTools)
library(tidyverse)
library(tidycensus)
library(purrr)
library(DT)
```

```{r gather_data}
### STEP 1. GATHER DATA FOR ANALYSIS

# a. Get Projects from LIMS
pi <- parseWebReport("https://nasis.sc.egov.usda.gov/NasisReportsWebSite/limsreport.aspx?report_name=get_project_from_NASISWebReport", args = list(p_mlrassoarea = params$mlrassoarea, p_fy = params$fiscalyear))

# b. Get MUKEYS from LIMS
polys <- parseWebReport('https://nasis.sc.egov.usda.gov/NasisReportsWebSite/limsreport.aspx?report_name=WEB-PROJECT_MUKEY_BY_GOAL_YEAR', args = list(msso = params$mlrassoarea, fy = params$fiscalyear, asym='%', proj='0'))

# c. Get Population Data from US Census
census_api_key("02f3fa45f183f3343f69234d1f329a2c3cb294c8")

us <- unique(fips_codes$state)[1:51]

totalpop <- map_df(us, function(x) {
  get_acs(geography = "county", variables = "B01003_001", 
          state = x)
})

# d. Get Spatial Data (can only handle one SSO at a time)
ssmu <- read_sf(dsn = paste0("D:/geodata/project_data/11", str_sub(params$mlrassoarea, -3, -1), "/RTSD_Region_11-", str_sub(params$mlrassoarea, -3, -1), "_FY19.gdb"), layer = "MUPOLYGON")

# e. Get custom CSV with SSO and Region 11 totals
tsso <- read.csv("c:/workspace2/impact_data.csv")
```

```{r clean_data}

### STEP 2. CLEAN DATA

## CLEAN MUKEY DATA FROM LIMS (polys)

# a. rename columns
names(polys) <- make.names(names(polys))

# b. remove description column (it's full of NAs anyway since we specifically asked not to return it in the above query)
polys$Description <- NULL

# c. make an index for removing unwanted project types
idx <- grep("EVAL|SDJR|MLRA", polys$Project.Type.Name)

# d. keep the only the rows matching the index
polys <- polys[idx, ]

# e. rename column 10 FIPS (there was a duplicate Area.Symbol name generated earlier)
colnames(polys)[10] <- "FIPS"

# f. make a new column of GEOID (state and county fips) using FIPS column and fips function
polys$GEOID <- paste0(fips(substr(polys$FIPS, 1,2)), str_sub(polys$FIPS, -3, -1))

# g. make an index for removing unwanted population data
idx3 <- grep(paste(polys$GEOID, collapse = "|"), totalpop$GEOID)

# h. keep only the rows matching the index
totalpop <- totalpop[idx3, ]

# i. convert to data.frame
totalpop <- as.data.frame(totalpop)

# j. join polys data and population data by GEOID
polypop <- left_join(polys, totalpop, by = "GEOID")

# k. sum the total population for each project
prpop <- aggregate(polypop$estimate, by = list(projectname = polypop$Project.Name), FUN = sum)

# l. rename column for joining later on
colnames(prpop)[2] <- "population"


## CLEAN SPATIAL DATA (ssmu)

# a. make an index for removing unwanted polygons
idx2 <- grep(paste(polys$mukey, collapse = "|"), ssmu$MUKEY)

# b. keep only the rows matching the index
ssmu <- ssmu[idx2, ]

# c. rename column for joining
colnames(ssmu)[3] <- "mukey"

# d. join project names with polygons in spatial data
ssmu_poly <- merge(ssmu, polys[, c("mukey", "Project.Name")], by = "mukey")

# e. remove geometry (we don't need it)
ssmu_poly$geometry <- NULL

# f. return a count of the polygons in each project
sm_ssmu_poly <- count(ssmu_poly, Project.Name)

# g. rename columns in the combined polygons data (ssmu) for joining later on
sm_ssmu_poly$projectname <- sm_ssmu_poly$Project.Name
sm_ssmu_poly$polygons <- sm_ssmu_poly$n
```

```{r combine_data}

### STEP 3. COMBINE PROJECT DATA TO SSO TOTALS DATA CALCULATE IMPACT SCORE AND DISPLAY RESULTS

# a. rename column in SSO totals (tsso) for joining
tsso$mlrassoarea <- as.character(tsso$Office.Area)

# b. join project data and totals
pd_tls <- merge(pi, tsso[, c("mlrassoarea", "Total.Area.SYMS", "Total.NATSYMS", "Total.Polygons", "Total.Acres", "total.population")], by = "mlrassoarea")
 
# c. join spatial data and project data together
sm_ssmu_poly <- merge(sm_ssmu_poly, prpop[, c("projectname","population")], by = "projectname")

# d. join totals data and sso data together
imp_d <- merge(pd_tls, sm_ssmu_poly[, c("projectname", "polygons", "population")], by = "projectname")

# e. calculate weighted impact factors
imp_d$im_asym <- (imp_d$n_areasymbol / imp_d$Total.Area.SYMS) * 20
imp_d$im_natsym <- (imp_d$n_nationalmusym / imp_d$Total.NATSYMS) * 20
imp_d$im_poly <- ( imp_d$polygons / imp_d$Total.Polygons) * 20
imp_d$im_acres <- (imp_d$acre_landcat / imp_d$Total.Acres) * 20
imp_d$im_popu <- (imp_d$population / imp_d$total.population) * 20

# f. sum the individual factors
imp_d$impact_score <- imp_d$im_asym + imp_d$im_natsym + imp_d$im_poly + imp_d$im_acres + imp_d$im_popu

# g. keep only the columns needed
final_d <- imp_d[, c(2, 1, 35,36,37,38,39,40)]

# h. round values for easier reading
final_d[, c(-1,-2)] <- round(final_d[, c(-1, -2)], 2)

# i. save file to .csv
write.csv(final_d, paste0(final_d$mlrassoarea[1], "_imp_score", ".csv"))

# j. make a DT data table
final_dt <- datatable(final_d)

# k. display table
final_dt
```
