---
title: "Saturated Hydraulic Conductivity in the Dillon Creek and Hall Creek Watersheds"
author: "John Hammerly"
date: "`r Sys.Date()`"
output: html_document
---
# Study Area

The study area of interest includes 2 similar watershed areas in southern Indiana. Observations and measurements were taken in the Dillon Creek watershed where a model was developed.  Hall Creek watershed is where the model was applied.  The extents of the watersheds mapped below, show the Dillon Creek Watershed matches the HUC12 unit.  The Hall Creek Watershed were the model was applied is a subset of the HUC12 Hall Creek Watershed.

```{r include = FALSE}

# load required libraries
library(sf)
library(ggplot2)
library(raster)
library(viridis)
library(knitr)
library(sf)
library(soilDB)
library(dplyr)
library(readxl)

# read study area shapefiles
dc <- st_read("D:/Cloud/Annum_Data/FY2019/Region 11/NO/INTERP/spatial_data/Dillon_HUC.shp")
dh <- st_read("D:/Cloud/Annum_Data/FY2019/Region 11/NO/INTERP/spatial_data/Dillon_Hall_HUC.shp")
hc <- st_read("D:/Cloud/Annum_Data/FY2019/Region 11/NO/INTERP/Indiana_DSM_Project/HUC12_Hall_Creek/watershed_30m.shp")

shc <- st_read("D:/Cloud/Annum_Data/FY2019/Region 11/NO/INTERP/spatial_data/SSURGO/spatial/soilmu_a_aoi.shp", stringsAsFactors = FALSE)
sdc <- st_read("D:/Cloud/Annum_Data/FY2019/Region 11/NO/INTERP/spatial_data/D_Creek_SSURGO/spatial/soilmu_a_aoi.shp")

```


```{r fig.height=8, fig.width=10, message=FALSE, warning=FALSE, echo = FALSE, message = FALSE}

# plot study area
ggplot() +
  geom_sf(data = dh) +
  geom_sf(data = hc, fill = "blue4", alpha = .75) +
  geom_sf(data = dc, fill = "blue4", alpha = .75) +
  geom_sf_label(data = dh, aes(label = dh$NAME)) +
  geom_sf_label(data = hc, aes(label = "Hall Creek Subset"), nudge_y = .01) +
  theme_minimal() +
  labs(x = "Longitude", y = "Latitude")
```

# Dillon Creek

Amoozemeter measurements were collected in pasture and forest catchments.  Sites and pedons are in NASIS.  Amoozemeter measurements are not.  Adding them would require a permission change to Region 11, or a willing data entry staff member from Region 6.

## Catchment Locations

The forest catchment is located just outside the watershed.

```{r fig.height=5, fig.width=10, message=FALSE, warning=FALSE, echo = FALSE, }

# fetch NASIS pedons from study area
p <- fetchNASIS()

s <- site(p)

s <- s %>% mutate(chmnt = "pasture")

s[9:13, 92] <- "forest"

pc <- p[1:8]

sp <- site(pc)

fc <- p[9:13]

fp <- site(fc)

# plot pedon locations
ggplot(dc) +
  geom_sf() +
  theme_minimal() +
  geom_point(data = s, mapping = aes(x = x_std, y = y_std, color = chmnt)) +
  labs(x = "Longitude", y = "Latitude", color = "Catchment")
```

## Pasture Catchment

### NASIS Pedon Profiles

These pedons have correlated taxonomic classifications.

```{r fig.height=5, fig.width=10, message=FALSE, warning=FALSE, echo = FALSE}

# plot pedon profiles
plot(pc, label = "site_id", alt.label = "taxonname", alt.label.col = "white")

```

### NASIS Pedon Locations

```{r fig.height=5, fig.width=10, message=FALSE, warning=FALSE, echo = FALSE}

# plot pedon locations
ggplot(dc) +
  geom_sf() +
  theme_minimal() +
  geom_point(data = sp, mapping = aes(x = x_std, y = y_std, color = taxonname)) +
  labs(x = "Longitude", y = "Latitude", color = "NASIS Series") +
  coord_sf(x = c(min(sp$x_std)-.002, max(sp$x_std)+.002), y = c(min(sp$y_std)-.002, max(sp$y_std)+.002)) +
  geom_text(data = sp, mapping = aes(x = x_std, y = y_std, label = c(1:8)), nudge_x =.0001)
  
```

### Amoozemeter Ksat Data

Measurement data includes a mean ksat and standard deviation for each of the 5 repetitions at each site.  Comparing the series in the table below to the NASIS pedons, you can see discrepancies.  Revisiting the classification of these soils and updating in NASIS and this table is needed.  The table below is a cleaned version of the summary of raw data.

```{r message=FALSE, warning=FALSE, echo = FALSE}

# Load up pasture ksat summary data and clean it up 

pksat <- read_excel("D:/Cloud/Annum_Data/FY2019/Region 11/NO/INTERP/Indiana_DSM_Project/Pasture_Catchment/Soil_Pits/Ksat_Data/Ksat_Soil_Moisture_Pits.xls", skip = 2)

# reformat field labels, removing dash
pksat$`Field Label` <- gsub("-0", "0", pksat$`Field Label`)

# move last digit of field label to a new layer column
pksat <- pksat %>% mutate(layer = substr(pksat$`Field Label`, 13,13))

# keep only the site_id from field label, make new site_id column, convert cm/hr ksat to micrometers/second
pksat <- pksat %>% mutate(site_id = substr(pksat$`Field Label`, 1, 11), um_ksat = round(pksat$`cm hr-1`*2.778, 4), position = pksat$`Soil Series...5`)

# replace series names in position column with position
pksat[c(1,4,7,10,13,15,19,22),11] <- pksat[c(2,5,8,11,14,16,20,23),11]

# replace NAs with position
pksat[c(3,6,9,12,17,18,21,24), 11] <- pksat[c(2,5,8,11,16,16,20,23),11]

# change position to lowercase
pksat <- pksat %>% mutate_at(11, list(position = tolower))

# remove spaces in position
pksat$position <- gsub(" ", "", pksat$position)

# rename columns column
pksat <- pksat %>% rename("taxonname" = `Soil Series...5`, "ssurgo_taxonname" = `Soil Series...6`, "taxon_class" = ...7, "horizon" = Horizon, "depb" = `Depth (cm)`)

# copy series to other rows
pksat[c(2,3,5,6,8,9,11,12,14,16,17,18,20,21,23,24),5] <- pksat[c(1,1,4,4,7,7,10,10,13,15,15,15,19,19,22,22),5]

# copy ssurgo series to other rows
pksat[c(2,3,5,6,8,9,11,12,14,16,17,18,20,21,23,24),6] <- pksat[c(1,1,4,4,7,7,10,10,13,15,15,15,19,19,22,22),6]

# copy taxon class to other rows
pksat[c(2,3,5,6,8,9,11,12,14,16,17,18,20,21,23,24),7] <- pksat[c(1,1,4,4,7,7,10,10,13,15,15,15,19,19,22,22),7]

# add top depth column
pksat <- pksat %>% group_by(site_id) %>% mutate(dept = lag(depb))

# add zero depths for surface
pksat$dept[is.na(pksat$dept)] <- 0

# reorder columns
pksat <- pksat %>% select(site_id, layer, horizon, dept, depb, um_ksat, taxonname, ssurgo_taxonname, taxon_class, position)

# add catchment column
pksat <- pksat %>% mutate(catchment = "pasture")

# show table
kable(pksat)

```

### Amoozemeter Ksat Plots

Not all soil layers described had ksat measurements taken.  The plots below are of only the measured horizons.

```{r message=FALSE, warning=FALSE, echo = FALSE}

# calculate center of layers
pksat <- pksat %>% mutate(mdep = (depb + dept)/2)

# plot ksat vs. depth by pedon
ggplot(pksat) +
  geom_text(
    aes(
      y     = pksat$um_ksat,
      x     = pksat$mdep,
      label = pksat$horizon
    ),
    size    = 4,
    hjust   = 0,
    nudge_y = .1
  ) +
  coord_flip() +
  scale_x_reverse() +
  geom_step(
    direction = "vh", aes(y = pksat$um_ksat, x = pksat$dept)
  ) +
  geom_step(
    aes(
      y = pksat$um_ksat,
      x = pksat$depb)
  ) +
  geom_point(
    aes(
      y = pksat$um_ksat,
      x = pksat$mdep
      ),
      color = "red"
  ) +
  facet_wrap(~ pksat$site_id) +
  labs(
    title = "Mean Values by Pedon",
    y = expression(
         paste(K[sat], " (micrometers/second)")
        ),
    x = "Depth (cm)"
  ) +
  expand_limits(y = pksat$um_ksat * 10
  ) +
  scale_y_log10(breaks = c(0, 0.01, 0.1, 1, 10, 100), minor_breaks = waiver(), labels = waiver())
```

## Forest Catchment

### NASIS Pedon Profiles

These pedons taxonomic classification is still "sampled as", and should be updated.

```{r fig.height=5, fig.width=10, message=FALSE, warning=FALSE, echo = FALSE}

# plot pedon profiles
plot(fc, label = "site_id", alt.label = "taxonname", alt.label.col = "white")
```

### NASIS Pedon Locations

These measurements were taken just beyond the boundary of the watershed.

```{r fig.height=5, fig.width=10, message=FALSE, warning=FALSE, echo = FALSE}

# plot pedon locations
ggplot(dc) +
  geom_sf() +
  theme_minimal() +
  geom_point(data = fp, mapping = aes(x = x_std, y = y_std, color = taxonname)) +
  labs(x = "Longitude", y = "Latitude", color = "Series") +
  coord_sf(x = c(min(fp$x_std)-.002, max(fp$x_std)+.002), y = c(min(fp$y_std)-.002, max(fp$y_std)+.002)) +
  geom_text(data = fp, mapping = aes(x = x_std, y = y_std, label = c(1:5)), nudge_y =-.0001, nudge_x =.00005)
```

### Amoozemeter Ksat Data

```{r message=FALSE, warning=FALSE, echo = FALSE}

# Load up forest ksat summary data and clean it up 
fksat <- read_excel("D:/Cloud/Annum_Data/FY2019/Region 11/NO/INTERP/Indiana_DSM_Project/Forest_Catchment/Soil_Pits/Ksat_Data/Ksat_Soil_Moisture_Pits.xls", skip = 1)

# move position from field label column to new column
fksat <- fksat %>% mutate("position" = `Field Label`, um_ksat = round(fksat$`cm hr-1`*2.778, 4))

# remove field labels
fksat[c(1,5,9,12,15),8] <- fksat[c(2,6,10,13,16), 8]

# remove position from field label column
fksat[c(2,6,10,13,16), 1] <- fksat[c(1,5,9,12,15),1]

# copy field labels to other rows
fksat[c(3,4,7,8,11,14,17), 1] <- fksat[c(1,1,5,5,9,12,15),1]

# copy position to other rows
fksat[c(3,4,7,8,11,14,17), 8] <- fksat[c(1,1,5,5,9,12,15),8]

# change position to lowercase
fksat <- fksat %>% mutate_at(8, list(position = tolower))

# remove spaces in position
fksat$position <- gsub(" ", "", fksat$position)

# make new layer column
fksat <- fksat %>% group_by(`Field Label`) %>% mutate("layer" = as.character(row_number()))

# rename columns column
fksat <- fksat %>% rename("site_id" = `Field Label`, "taxonname" = `Soil Series...5`, "ssurgo_taxonname" = `Soil Series...6`, "taxon_class" = `Field Taxonomic Classification`, "horizon" = Horizon, "depb" = `Depth (cm)`)

# add top depth column
fksat <- fksat %>% group_by(site_id) %>% mutate(dept = lag(depb))

# add zero depths for surface
fksat$dept[is.na(fksat$dept)] <- 0

# copy series to other rows
fksat[c(2:4,6:8,10:11,13:14,16:17),5] <- fksat[c(1,1,1,5,5,5,9,9,12,12,15,15),5]

# copy ssurgo series to other rows
fksat[c(2:4,6:8,10:11,13:14,16:17),6] <- fksat[c(1,1,1,5,5,5,9,9,12,12,15,15),6]

# copy taxon class to other rows
fksat[c(2:4,6:8,10:11,13:14,16:17),7] <- fksat[c(1,1,1,5,5,5,9,9,12,12,15,15),7]

# ungroup
fksat <- ungroup(fksat)

# reorder columns
fksat <- fksat %>% dplyr::select(site_id, layer, horizon, dept, depb, um_ksat, taxonname, ssurgo_taxonname, taxon_class, position)

# add catchment column
fksat <- fksat %>% mutate(catchment = "forest")

# show table
kable(fksat)
```

### Amoozemeter Ksat Plots

```{r message=FALSE, warning=FALSE, echo = FALSE}

# calculate center of layers
fksat <- fksat %>% mutate(mdep = (depb + dept)/2)

# plot ksat vs. depth by pedon
ggplot(fksat) +
  geom_text(
    aes(
      y     = fksat$um_ksat,
      x     = fksat$mdep,
      label = fksat$horizon
    ),
    size    = 4,
    hjust   = 0,
    nudge_y = .1
  ) +
  coord_flip() +
  scale_x_reverse() +
  geom_step(
    direction = "vh", aes(y = fksat$um_ksat, x = fksat$dept)
  ) +
  geom_step(
    aes(
      y = fksat$um_ksat,
      x = fksat$depb)
  ) +
  geom_point(
    aes(
      y = fksat$um_ksat,
      x = fksat$mdep
      ),
      color = "red"
  ) +
  facet_wrap(~ fksat$site_id) +
  labs(
    title = "Mean Values by Pedon",
    y = expression(
         paste(K[sat], " (micrometers/second)")
        ),
    x = "Depth (cm)"
  ) +
  expand_limits(y = fksat$um_ksat * 10
  ) +
  scale_y_log10(breaks = c(0, 0.01, 0.1, 1, 10, 100), minor_breaks = waiver(), labels = waiver())
```

```{r echo = FALSE}

# combine pasture and forest datasets
aksat <- bind_rows(pksat, fksat)

```

### SSURGO Ksat

Ksat data from SSURGO indicates little variability at the surface, and greater variability at the bottoms of the soil profiles.

```{r echo=FALSE, fig.keep = "hold", out.width = "50%", message=FALSE, warning=FALSE}

# Ksat for Dillon Creek
cpnts <- fetchNASIS(from = "components")

mpunts <- get_mapunit_from_NASIS()

corr <- get_component_correlation_data_from_NASIS_db()

corr <- corr %>% mutate(lmapunitiid = as.character(lmapunitiid))

sdcmu <- sdc %>% left_join(corr, by = c("MUKEY" = "lmapunitiid"))

sdcmudmu <- sdcmu %>% left_join(cpnts@site, by = "dmuiid")

cpntshz <- cpnts@horizons %>% mutate(coiid = as.character(coiid))

s_dmu_co <- sdcmudmu %>% left_join(cpntshz, by = "coiid")

sdc_ksat1 <- s_dmu_co %>% filter(hzdept_r == 0) %>%  group_by(nationalmusym) %>% summarize(surf_ksat = round(weighted.mean(ksat_r, w = comppct_r), 2))

# plot ksat values
# plot(sdc_ksat1["surf_ksat"], breaks = c(0,10,20,30,40,50,60,70,80,90,100), pal =rev(viridis(10)), main = "Ksat, top layer - even scale")

sdc_ksat1_t <- st_drop_geometry(sdc_ksat1)



# kable(sdc_ksat1_t)

# plot using ksat class breaks
plot(sdc_ksat1["surf_ksat"], pal = rev(viridis(5)), logz = TRUE, breaks = c(-3, -2, -1, 0, 1, 2), at = c(-3, -2, -1, 0, 1, 2), main = "Ksat, top layer - log scale")

sdc_ksat3 <- s_dmu_co %>% group_by(nationalmusym, compname, comppct_r) %>% top_n(1, hzdepb_r) %>%  group_by(nationalmusym) %>% summarize(bot_ksat = round(weighted.mean(ksat_r, w = comppct_r, na.rm = TRUE), 2))

# plot ksat values
# plot(sdc_ksat3["bot_ksat"], pal =  rev(viridis(10)), main = "Ksat, bottom layer, - even scale")

sdc_ksat3_t <- st_drop_geometry(sdc_ksat3)

bxpgrp1 <- mutate(sdc_ksat1_t, layer = "top", "ksat" = surf_ksat)
bxpgrp3 <- mutate(sdc_ksat3_t, layer = "bottom", "ksat" = bot_ksat)


bxp <- bind_rows(bxpgrp1, bxpgrp3)

# kable(sdc_ksat3_t)

# plot using ksat class breaks
plot(sdc_ksat3["bot_ksat"], pal =  rev(viridis(5)),logz = TRUE, breaks = c(-3, -2, -1, 0, 1, 2), at = c(-3, -2, -1, 0, 1, 2), main = "Ksat, bottom layer - log scale")
```

```{r echo = FALSE}
ggplot(bxp) +
stat_boxplot(data = bxp, aes(y = ksat, x = layer)) +
scale_y_log10() +
  labs(x = "Layer")

summary(bxp)
```

```{r fig.height=15, fig.width=10, warning = FALSE, echo = FALSE, message = FALSE}
# calculate center of layers

s_dmu_co2 <- st_drop_geometry(s_dmu_co)

s_dmu_co2 <- s_dmu_co2 %>% mutate(mdep = (hzdepb_r + hzdept_r)/2)

s_dmu_co2 <- s_dmu_co2 %>% filter(!is.na(ksat_r))

# plot ksat vs. depth by pedon
ggplot(s_dmu_co2) +
  geom_text(
    aes(
      y     = s_dmu_co2$ksat_r,
      x     = s_dmu_co2$mdep,
      label = s_dmu_co2$hzname
    ),
    size    = 4,
    hjust   = 0,
    nudge_y = .1,
    alpha = .01
  ) +
  coord_flip() +
  scale_x_reverse() +
  geom_step(
    direction = "vh", aes(y = s_dmu_co2$ksat_r, x = s_dmu_co2$hzdept_r)
  ) +
  geom_step(
    aes(
      y = s_dmu_co2$ksat_r,
      x = s_dmu_co2$hzdepb_r)
  ) +
  geom_point(
    aes(
      y = s_dmu_co2$ksat_r,
      x = s_dmu_co2$mdep
      ),
      color = "red"
  ) +
  facet_wrap(~ s_dmu_co2$compname + s_dmu_co2$compkind) +
  labs(
    title = "Mean Values by Component",
    y = expression(
         paste(K[sat], " (micrometers/second)")
        ),
    x = "Depth (cm)"
  ) +
  expand_limits(y = s_dmu_co2$ksat_r * 10
  ) +
  scale_y_log10(breaks = c(0, 0.01, 0.1, 1, 10, 100), minor_breaks = waiver(), labels = c(0, 0.01, 0.1, 1, 10, 100))
```


# Hall Creek Watershed Subset

### SSURGO Ksat

```{r echo=FALSE, fig.keep= "hold", out.width="50%", message=FALSE, warning=FALSE}
shcmu <- shc %>% left_join(corr, by = c("MUKEY" = "lmapunitiid"))

shcmudmu <- shcmu %>% left_join(cpnts@site, by = "dmuiid")

cpntshz_hc <- cpnts@horizons %>% mutate(coiid = as.character(coiid))

s_dmu_co_hc <- shcmudmu %>% left_join(cpntshz_hc, by = "coiid")

shc_ksat1 <- s_dmu_co_hc %>% filter(hzdept_r == 0) %>%  group_by(nationalmusym) %>% summarize(surf_ksat = round(weighted.mean(ksat_r, w = comppct_r), 2))

# plot ksat values
# plot(shc_ksat1["surf_ksat"], breaks = c(0,10,20,30,40,50,60,70,80,90,100), pal =  rev(viridis(10)), main = "Ksat, top layer")

shc_ksat1_t <- st_drop_geometry(shc_ksat1)

# kable(shc_ksat1_t)

# plot using ksat class breaks
plot(shc_ksat1["surf_ksat"], pal =  rev(viridis(5)),logz = TRUE, breaks = c(-3, -2, -1, 0, 1, 2), at = c(-3, -2, -1, 0, 1, 2), main = "Ksat, top layer - log scale")


shc_ksat3 <- s_dmu_co_hc %>% group_by(nationalmusym, compname, comppct_r) %>% top_n(1, hzdepb_r) %>%  group_by(nationalmusym) %>% summarize(bot_ksat = round(weighted.mean(ksat_r, w = comppct_r, na.rm = TRUE), 2))

# plot ksat values
# plot(shc_ksat3["bot_ksat"], breaks = c(0,10,20,30,40,50,60,70,80,90,100), pal =  rev(viridis(10)), main = "Ksat, bottom layer - even scale")

shc_ksat3_t <- st_drop_geometry(shc_ksat3)

bxhpgrp1 <- mutate(shc_ksat1_t, layer = "top", "ksat" = surf_ksat)
bxhpgrp3 <- mutate(shc_ksat3_t, layer = "bottom", "ksat" = bot_ksat)


hbxp <- bind_rows(bxhpgrp1, bxhpgrp3)



# kable(shc_ksat3_t)

# plot using ksat class breaks
plot(shc_ksat3["bot_ksat"], pal =  rev(viridis(5)),logz = TRUE, breaks = c(-3, -2, -1, 0, 1, 2), at = c(-3, -2, -1, 0, 1, 2), main = "Ksat, bottom layer - log scale")
```

```{r, echo = FALSE}

ggplot(hbxp) +
stat_boxplot(data = hbxp, aes(y = ksat, x = layer)) +
scale_y_log10() +
  labs(x = "Layer")

summary(hbxp)

```

```{r fig.height=8, fig.width=10, message=FALSE, warning=FALSE, echo = FALSE}

# read ksat layers and convert to micrometers per second
kt <- raster("D:/workspace/hcr/hcr_ksat_l1")*2.778
km <- raster("D:/workspace/hcr/hcr_ksat_l2")*2.778
kb <- raster("D:/workspace/hcr/hcr_ksat_l3")*2.778
ka <- raster("D:/workspace/hcr/hcr_ksat_l123")*2.778

# stack layers together
d <- stack(ka,kt,km,kb)

```

## Modeled Ksat Layers

### Even Scale

```{r echo=FALSE, fig.height=8, fig.width=10, message=FALSE, warning=FALSE}

# set breaks, based on min/max values
brks <- seq(0,2.75, by=0.01)

# one less for colorscale
nb <- length(brks)-1

# build colors
cols <- rev(viridis(nb))

plot(d, breaks = brks, col = cols, axis.args = list(at = c(0, 0.25, 0.5, .75, 1, 1.25, 1.5, 1.75, 2, 2.25, 2.5, 2.75), cex.axis = 0.75, labels = c("0.00", "0.25", "0.50", "0.75", "1.00", "1.25", "1.50", "1.75", "2.00", "2.25", "2.50", "2.75")), zlim = c(0, 2.75), legend.width = 2, legend.shrink=0.75, main = c(expression("K"["sat"]~"all layers ("*mu*m*"/sec)"), expression("K"["sat"]~"top layer ("*mu*m*"/sec)"), expression("K"["sat"]~"middle layer ("*mu*m*"/sec)"), expression("K"["sat"]~"bottom layer ("*mu*m*"/sec)")))

```

### Ksat Classes

```{r echo=FALSE, fig.height=8, fig.width=10, message=FALSE, warning=FALSE}

brks2 <- c(.001, 0.01, 0.1, 1, 10, 100)

nb2 <- length(brks2)-1

cols2 <- rev(viridis(nb2))

plot(d, legend = FALSE, breaks = brks2, col = cols2, main = c(expression("K"["sat"]~"all layers"), expression("K"["sat"]~"top layer"), expression("K"["sat"]~"middle layer"), expression("K"["sat"]~"bottom layer ")))
legend("bottom", legend = c("Very Low","Low", "Moderately Low", "Moderately High", "High"), fill = cols2)
```

### Boxplots

```{r fig.height=8, fig.width=10, message=FALSE, warning=FALSE, echo = FALSE}

# generate boxplots

kat <- as.data.frame(ka)
kat <- kat %>% rename("ksat" = hcr_ksat_l123) %>% mutate("layer" = "all")
ktt <- as.data.frame(kt)
ktt <- ktt %>% rename("ksat" = hcr_ksat_l1) %>% mutate("layer" = "top")
kmt <- as.data.frame(km)
kmt <- kmt %>% rename("ksat" = hcr_ksat_l2) %>% mutate("layer" = "mid")
kbt <- as.data.frame(kb)
kbt <- kbt %>% rename("ksat" = hcr_ksat_l3) %>% mutate("layer" = "bottom")

akt <- bind_rows(kat, ktt, kmt, kbt)

ggplot(akt) +
stat_boxplot(aes(y = ksat , x = layer)) +
scale_y_log10() +
  labs(x = "Layer")

```

## Quantiles Table

```{r echo = FALSE, fig.height=8, fig.width=10, message=FALSE, warning=FALSE}

# generate quantile table
rqt <- as.data.frame(quantile(d))

# show table
kable(rqt)
```

